Leader functions and qualities across 59 cultures: an exploratory analysis

### 2019-12-03 leadership_across_cultures_contexts.Rmd
Other edits in discussion sections, etc.

Across a broad range of cultures and social contexts leaders are nearly universally high status males who are described as knowledgeable or intelligent and who are experienced and have a demonstrated history of previous accomplishments or special skills. They most frequently resolve conflicts within their respective groups while also providing important social functions, organizing cooperation among followers, providing counsel or input in followers' affairs, and distributing resources. @oc06-034 (p. 236) in describing kin group leaders among the Iban, small-scale highland horticulturalists primarily in Borneo and Sarawak, provides example of this suite of traits in leaders:

@@

We performed hierarchical cluster analysis on our leader quality and leader function variables to determine if these variables formed theoretically meaningful categories. Variables were clustered together if they were positively correlated. Leader qualities formed two top-level clusters, which we identify as a *Material, social, and reproductive success* cluster, and a *Social orientation* cluster. See Figure \@ref(fig:cluster)A. Several of traits often co-occur. Among the Trobriands, horticulturalists in Papua New Guinea, @ol06-037 (p. 47 ) describes relationships between polygnyous marriages, increased wealth, and widespread social contacts:

@@

The larger organization subcluster included variables such as *Organize cooperation*, *Ritual functions*, and *Military command*. This subcluster aligns with the many evolutionary theories of leadership that emphasize the role of leaders in organizing collective actions [e.g., @hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; for review, see @garfield_evolutionary_2019]. In the ethnographic record such functions includes controlling and facilitating economic productivity and ensuring collective actions do not overlly conflict with individual or household-level goals, such as among the Tukano of the Braziallian Amazon, where @sq19-018 (p. 127) notes:

@@

The strong support for material and social benefits associated with leadership (Figure \@ref(fig:varsupportcb)) and the colinearity of measures associated with material wealth, social capital, and reproductive success (Figure \@ref(fig:cluster)A), suggests processes of leadership and followership might be strong drivers of the emergence and intergenerational transmission of inequality [@bowles_emergence_2010; @mattison_evolution_2016]. Various sources of material wealth and strategic or institutionalized processes of nepotism are widespread cross-culturally, including among many societies often described as "egalitarian." Theses results suggests anthropologists and economists investigating the emergence and evolution of inequality, including material, social, and reproductive dimensions, should incorporates and pay close attention to group processes and social strategies associated with leadership and followership.

@@

This study is based on the ethnographic record. Our methods are designed to capture a broad range of ethnographic descriptions from a culturally representative, systematic sample. Nonetheless, these results are conditional upon the content ethnographers chose to discuss and the way they chose to discuss it. The ethnographic record is certainly an imperfect source of information on human cultural diversity and hence a limitation of the current results. We, however, view the existence of the ethnographic record and the availability and functionality of the eHRAF as an enormous benefit for comparative research, given many of the populations represented in these data no longer exist or live as they did during the time of much ethnographic description. Within these ethnographic descriptions there is a general male bias, which could reflect a male bias in community leadership or the more systemic male bias in the ethnographic record. These data are also biased towards cultural models (rather than descriptions of observed behavior) and community leadership; descriptions of kin group and residential group leadership are less frequent. The high proportion of text records discussing political group leaders could be attributable to our broad search strategy or the content of the ethnographic record of "leadership" more generally.

>>>>

Across a broad range of cultures and social contexts leaders are nearly universally high status males who are described as knowledgeable or intelligent and who are experienced and have a demonstrated history of previous accomplishments or special skills. They most frequently resolve conflicts within their respective groups while also providing important social functions, organizing cooperation among followers, providing counsel or input in followers' affairs, and distributing resources. @oc06-034 (p. 236) in describing kin group leaders among the Iban, small-scale highland horticulturalists primarily in Borneo and Sarawak, provides example of this suite of traits in leaders (which also include elements of prestige-based and charismatic leadership discussed above):

@@

We performed hierarchical cluster analysis on our leader quality and leader function variables to determine if these variables formed theoretically meaningful categories. Variables were clustered together if they were positively correlated. Leader qualities formed two top-level clusters, which we identify as a *Material, social, and reproductive success* cluster, and a *Social orientation* cluster. See Figure \@ref(fig:cluster)A. Several of traits often co-occur. Among the Trobriands, horticulturalists in Papua New Guinea, @ol06-037 (p. 47 ) describes relationships between polygynous marriages, increased wealth, and widespread social contacts:

@@

The larger organization subcluster included variables such as *Organize cooperation*, *Ritual functions*, and *Military command*. This subcluster aligns with the many evolutionary theories of leadership that emphasize the role of leaders in organizing collective actions [e.g., @hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; for review, see @garfield_evolutionary_2019]. In the ethnographic record such functions includes controlling and facilitating economic productivity and ensuring collective actions do not overly conflict with individual or household-level goals, such as among the Tukano of the Brazilian Amazon, where @sq19-018 (p. 127) notes:

@@

The strong support for material and social benefits associated with leadership (Figure \@ref(fig:varsupportcb)) and the colinearity of measures associated with material wealth, social capital, and reproductive success (Figure \@ref(fig:cluster)A), suggests processes of leadership and followership might be strong drivers of the emergence and intergenerational transmission of inequality [@bowles_emergence_2010; @mattison_evolution_2016]. Various sources of material wealth and strategic or institutionalized processes of nepotism are widespread cross-culturally, including among many societies often described as "egalitarian." Theses results suggests anthropologists and economists investigating the emergence and evolution of inequality, including material, social, and reproductive dimensions, should incorporate and pay close attention to group processes and social strategies associated with leadership and followership.

@@

This study is based on the ethnographic record. Our methods are designed to capture a broad range of ethnographic descriptions from a culturally representative, systematic sample. Nonetheless, these results are conditional upon the content ethnographers chose to discuss and the way they chose to discuss it. The ethnographic record is certainly an imperfect source of information on human cultural diversity and hence a limitation of the current results. We, however, view the existence of the ethnographic record and the availability and functionality of the eHRAF as an enormous benefit for comparative research, given many of the populations represented in these data no longer exist or live as they did during the time of much ethnographic description. Within these ethnographic descriptions there is a general male bias, which could reflect a male bias in community leadership or the more systemic male bias in the ethnographic record [see @rosaldo_woman_1974; @mukhopadhyay_anthropological_1988]. These data are also biased towards cultural models (rather than descriptions of observed behavior) and community leadership; descriptions of kin group and residential group leadership are less frequent. The high proportion of text records discussing political group leaders could be attributable to our broad search strategy or the content of the ethnographic record of "leadership" more generally.

Another limitation lies in our ability to interpret the absence of evidence. We are confident that the cross-culturally frequent measures discussed here and strongly represented in the ethnographic record and hence very likely to represent broad cross-cultural patterns in human behavior and culture. We cannot strongly conclude, however, that the relatively infrequent measures identified here are truly infrequent dimensions of human behavior and culture, only that they are relatively less frequent in the ethnographic record. Nonetheless, results reported here provide a broad, comprehensive, and cross-cultural framework for future empirical and theoretical work to draw on.
